\documentclass[]{article}

\usepackage{csquotes}
\usepackage{listings}

%opening
\title{Constructing a powerful and flexible computing cluster using Raspberry Pi computers}
\author{Camilo Correa Restrepo, Universidad EAFIT}

\begin{document}

\maketitle

\section{Introduction}
Not too long ago, creating a cluster of computers, using affordable and commercially available hardware,
    that was suitably portable for education and durable enough to withstand the harshest conditions a classroom
    could offer was a task not very well suited for the faint of heart, specially because of the fact that
    wrestling for hours on end with unreliable software would probably wreak havoc on one's mind and hardware.
    However, as time has gone on, and with the advent of small, reasonably fast and extremely cheap computers, this idea
    no longer is a dream of many in the academic community, but rather, an interesting reality that's ripe for
    exploration. As such, we will now embark on the journey of constructing our very own cluster, with the quite
    sincerely amazing Raspberry Pi. With that in mind, the goal of this project is, evidently, to create a small
    and very easy to dismantle cluster that is very easy to set up and get up and running for educational
    purposes (for after all, you will not be able to do very complex computations on this hardware). This cluster will be very easily 
    customizable, easy to manage and maintain, and will behave exactly like a real world high performance computer would on a much smaller 
    scale. \\

    Despite the simple nature of this project, I will, however, assume you have a basic knowledge and experience 
    using UNIX systems. We will set out to achieve a very simple goal to channel our efforts, as, without it,
    we would go nowhere. We will work towards making this system as robust and as well set up as we can, in order to prepare it to run one very special application, John the Ripper, that bridges the gap between the world of High Performance Computing and the world of Cyber Security. Though this might not even make sense, it is imperative that you understand, going into this, that the true purpose of this is to enable whoever may read it to acquire and absorb a myriad of skills that will allow them to continue enjoying and exploring this daring and fantastic new world with effectiveness and preparedness.

    \newpage
    \tableofcontents

    \newpage
    \section{What is a Raspberry Pi and why are we using it?}
    It is absolutely impossible to talk about low cost and low consumption devices without
    mentioning a raspberry pi, it is an absolutely phenomenal device, that carries has propelled
    forth a true paradigm shift in the way we look at embedded devices, and it is has also proven
    itself to be a truly spectacular way to teach the ways of computing to others. Thus, let me first begin by 
    presenting a quote from the Raspberry Pi Foundation, that presents what a 
    raspberry pi is better than I ever could:
    \begin{displayquote}
    ``The Raspberry Pi is a credit-card sized computer that plugs into your TV and a keyboard. It is a capable 
    little computer which can be used in electronics projects, and for many of the things that your desktop PC 
    does, like spreadsheets, word-processing, browsing the internet and games. It also plays high-definition 
    video. We want to see it being used by kids all over the world to learn 
    programming.''\footnote{https://www.raspberrypi.org/help/faqs/}
    \end{displayquote}

    As you can now, hopefully, imagine, these little devices are a true marvel of modern engineering,
    as they pack within a truly great ``bang for your buck'' in terms of capabilities, and are an 
    absolutely perfect testbed for large engineering solutions.
    \section{What you will need}
    In order to construct  your very own cluster, you will need, at the very least, the following components:
    \begin{itemize}
        \item 2 or more Raspberry Pi computers
            \item A micro-USB cable for every Raspberry Pi
                \item An Internet router with an integrated switch or a network switch
                    \item Enough Cat-5 cables to connect all computers to the network
                        \item SD cards for every Raspberry Pi of at least 4GB of space (ideally 8GB or larger).
                            \item A USB flash drive of at least 4GB of space (ideally 8GB or larger).
                                \item A USB keyboard
                                    \item An HDMI-capable screen
                                        \item A computer with an SD card reader.
                                        \end{itemize}

                                        And additionally, while not strictly necessary, these additional components will make your life a lot 
                                        easier:

                                        \begin{itemize}
                                            \item An Internet connection (extremely recommended)
        \item An additional computer with ssh capabilities.
            \item Spare SD cards (should one fail or get damaged)
        \item Spare flash drives (to expand the internal storage of the pi's)
            \item Cases for your Raspberry Pis, as they might get damaged when you move them around or connect them 
                to power sources or network interfaces.
                    \item USB wall adapters, so you can hook up your Raspberry Pis to your wall outlets.
                    \end{itemize}

                    Please do keep in mind that these are the bare essentials to begin the construction of the system; thus, to 
                    scale this system, all you'll need to do is begin expanding the quantities of the materials involved, and 
                    performing all setup operations on every new component. \\

                    Additionally, if you wish to create a larger array of pis, do keep in mind that their combined heat output 
                    will no longer be trivial, and, thusly, it would be in your best interest to begin thinking of a spatial 
                    distribution that would allow your array to receive an adequate amount of airflow.

                    \section{Initial tasks}
                    Assuming you have procured all of the necessary components, let us, then, focus our efforts on the first 
                    tasks that must be completed before we can begin to use our shiny new cluster.

                    \subsection{Choosing your operating System}
                    The first and most important step of the process, without a doubt, is the operating system you will use on 
                    your devices, simply due to the fact that while they will all be based on the Linux kernel, they all have 
                    large and small differences that will either make your life easy, or hell on earth, so you must be quite 
                    sure of the OS you are going to marry. In my humble opinion, the best choice for a raspberry pi cluster is 
                    Raspbian, the Raspberry Pi foundation's operating system, which is a fork of the much revered and loved 
                    Debian GNU/Linux operating system. This however, does not in any way demerit all the other available OSs, as 
                    they all have their pros and cons, and it is absolutely true that you will run into a couple of headaches by 
                    using Raspbian. However, it does, as of the time this is being written, offer the greatest amount of 
                    community support and the greatest amount of pre-built packages, which will make getting the system up and 
                    running far easier\footnote{A word about Raspbian must be said, before going forward: quite recently
                        Raspbian switched from being based on Debian Wheezy to Debian Jessie, which, while fantastic for early 
                            adopters and those that crave the latest packages, is not ideal for what we'll be doing for our build, 
                                     simply due to the fact that it has some features that make it not desirable for us, for instance, not all 
                                         packages have been ported over, and probably quite a few are still quite buggy, additionaly, its kernel 
                                         version is 4.1, which is known to have a couple of stability issues, specially on hardware that isn't 
                                         x86.}.\\

                                         If, despite reading everything above, you still feel the desire to work with another OS, I would personally 
                                         recommend Arch Linux as the best alternative, but do keep in mind that you will need to do a lot of 
                                         compiling on your own, and, it goes without saying, this definitely isn't for the faint of heart.

                                         \subsection{Setting up your master SD card}
                                         Once you have chosen a suitable operating system for your device, the next step is to begin preparing the SD 
                                         card for your master node\footnote{The one that'll be in charge of controlling all the other nodes}. First, 
                                         you should download a copy of the disk image of the operating system you wish to run. For example, for 
                                         Raspbian, head to the Raspberry Pi foundation's website and find the .img for Raspbian. Once you are done 
                                         downloading, all you need to do then is do a simple \textit{dd} command to copy the disk image onto your 
                                         shiny new SD card.

                                         \textbf{I must however issue a word of caution to anyone using dd for the first time, or for anyone that is 
                                             inexperienced: \textit{dd} is, like most core UNIX utilities, extremely useful and extremely dangerous, as 
                                                 such, I must emphasize that you must exercise extreme caution using the command, as any mistake in the 
                                                 output target could result in the loss of data.}

                                                 \subsection{Physically setting up your cluster}
                                                 Now that the SD card is ready, it is now time to begin preparing all the physical components you will be 
                                                 using. Depending on how you've chosen to work, and how many nodes you will have in your network, it might be 
                                                 in your best interest to physically assemble your cluster before doing any work on it. If, however, you plan 
                                                 to operate a simple and small cluster, then you may want to avoid setting too many things up beforehand, and 
                                                 simply getting straight to work on the cluster.

                                                 No matter what you've chosen, you will absolutely need, for your first setup, a network connection, a 
                                                 monitor, and a keyboard. As soon as you have hooked all of these up go ahead and insert your SD card into 
                                                 it's slot and connect your power cord.

                                                 \textbf{Once again I must issue a word of warning for those working through these steps. Raspberry Pis, in 
                                                     theory are compatible with a very large amount of peripheral devices, but, that does not guarantee that the 
                                                         particular ones you will be using will be supported, especially if, for example you want to use an HDMI to 
                                                         VGA adapter, or a wireless keyboard or a WiFi dongle. Nonetheless, there are very many internet resources 
                                                         ready to point you in the right direction, and most, if not all of what you need is only a google search 
                                                         away.}

                                                         \section{Setting up your master node}
                                                         If everything up to this point went well, then you will probably be staring at a prompt asking for your user 
                                                         and password. If you went with Raspbian, then you need only enter \textit{pi} as your user and 
                                                         \textit{raspberry} as your password. You will now see that you are logged in to your primary raspberry pi, 
                                                         and you are now absolutely ready to begin the setup of your cluster.

                                                         \subsection{Getting the latest versions of the default packages}
                                                         Once you've logged in, it is now time to begin the very easy process of obtaining all of the newest and 
                                                         shiniest packages for your system, as it is always a good idea to have the latest and most updated version 
                                                         of the base system utilities. To those of you familiar with the Aptitude package manager, it will be a true 
                                                         breeze to do this process. For those of you that aren't, all you'll need to do is input the following two 
                                                         lines:
                                                         \begin{lstlisting}[language=bash]
                                                         $ sudo apt-get update
                                                         $ sudo apt-get upgrade
                                                         \end{lstlisting}

                                                         \subsection{Running the configuration script}
                                                         Next, you will need to enable the ssh server on your raspberry pi, so that you can connect to it remotely 
                                                         (should you desire to connect from another more comfortable computer). To enter the configuration script, 
                                                         enter the command:
                                                         \begin{lstlisting}[language=bash]
                                                         $ sudo raspi-config
                                                         \end{lstlisting}
                                                         Using your arrow keys, enter the advanced options and navigate onto the ssh section, then enable the service 
                                                         and exit the script. Once you are done, your raspberry pi will reboot and we will be ready to continue.

                                                         \subsection{Mounting your USB drive}
                                                         Every linux user knows that mounting a usb drive is a rather trivial task, however, for our needs we will 
                                                         need to have this drive automatically mounted as soon as our system begins, thus, we will need to edit our 
                                                         \textit{/etc/fstab} file so that it is automatically mounted. As a convention, it would be a good idea to 
                                                         mount the drive to a folder such as \textit{/share}.

                                                         \subsection{Installing all the needed utilities}
                                                         Now that you having enabled ssh, it is time to download all the utilities that we will require in order to 
                                                         mount our cluster. Input to the following command\footnote{As you will see in the following command and in 
                                                             several other commands, a backslash ($\backslash$) will indicate that the command continues but just on the 
                                                                 next line, basically, just imagine that you paste whatever begins on the next line exactly on the spot that 
                                                                 the backslash is.} to download everything we'll need:
                                                                 \begin{lstlisting}[language=bash]
                                                                 $ sudo apt-get install gfortran g++ nmap rpcbind\ 
                                                                 nfs-kernel-server nfs-common tcl8.5 tcl8.5-dev vim\ 
                                                                 tree htop libssl-dev m4 gawk libc6-dev libstdc++6 libnss3
                                                                 \end{lstlisting}
                                                                 Go ahead and make yourself a cup of tea while you wait, as this will probably take a while to download.

                                                                 \subsection{Preparing your NFS}
                                                                 If everything is working well, and your flash drive is working correctly, then all you'll need to do to make 
                                                                 it available to other computers is to create a rule for your mounted folder on your \textit{/etc/exports} 
                                                                 file\footnote{Some of you might consider it unusual that I more or less skipped this part. However, since 
                                                                     this is supposed to be an eyeopener to this wonderful world, I have decided that it is only fair that both 
                                                                         mounting the USB drive and the NFS server should be a an exercise for the reader, in order to test his 
                                                                         abilities to both problem solve, and to make things work. This is easily justified by the fact that things 
                                                                         can and will go wrong no matter how faithfully you follow the guide, so, please do understand that this is 
                                                                         for your own good.}. Make sure that the correct permissions are set for all user, and that they have access 
                                                                         to the full subtree of that drive. Next reboot your device and pay close attention to the boot-up messages, 
                                                                         since somewhere along the way, you'll first see how your drive is mounted and next, you'll see how the NFS 
                                                                         services restart and how they broadcast your file system. Make sure that this worked correctly before you 
                                                                         proceed as it is absolutely vital to most of the steps we will do going forward. \\

                                                                         Next, you'll need to set up a couple folders on your shared directory so that we can, going forward, fill it 
                                                                         up with interesting and powerful software and other things. So, go ahead and do the following:
                                                                         \begin{lstlisting}
                                                                         $ sudo mkdir /share/apps/ && sudo mkdir /share/modules/
                                                                         \end{lstlisting}

                                                                         \subsection{Setting up your environment modules}
                                                                         The time has finally come to begin inserting into our cluster some of the best and most powerful software 
                                                                         out there. First up to bat is Environment Modules, a simple little program that eats up TCL scripts and 
                                                                         allows you to set and unset your working environment just how you like it, so that you can use the exact 
                                                                         version you want of a program. In fact, you can modify every environment variable in your system with it, 
                                                                         and you can also quickly and safely undo your changes. \\

                                                                         \subsubsection{Overview of what will be done}
                                                                         In order to get our modules up, we will need to compile them, since, as of this writing, it is not available 
                                                                         in any of the public repositories for Raspberry Pi, and, even if it were, it is quite a good exercise to try 
                                                                         and get this program up and running yourself. Additionally, compiling it ourselves allows us to control, 
                                                                         with the utmost precision, how we want it to behave. Specifically, we don't want it to be installed in a 
                                                                         default directory, but rather in our \textit{/share/apps} directory, so we only need to install it once and 
                                                                         it will be forever available to all other nodes, and, furthermore, we will be able to configure it to look 
                                                                         for its modulefiles (more on this later) where we want it to, and nowhere else. \\

                                                                         Some of you will break out in a cold sweat just from the word "compile", but you'll quickly realize that as 
                                                                         with  most things in life, it'll only hurt the first time, and every time after that it'll only get better. 
                                                                         So, with that in mind let us begin.

                                                                         \subsubsection{Obtaining the Software}
                                                                         Since we won't be using our trusty if a bit temperamental friend Aptitude, we'll need to nab ourselves the 
                                                                         compressed package for Modules, which is easily obtainable from it's website using a simple \textit{wget} command and 
                                                                         a \textit{tar} command in order to access all the contents of the downloaded file.
                                                                         \begin{lstlisting}
                                                                         $ cd
                                                                         $ wget http://sourceforge.net/projects/modules/files\
                                                                             /Modules/modules-3.2.10/modules-3.2.10.tar.gz/
                                                                             $ tar xf modules-3.2.10.tar.gz
                                                                             \end{lstlisting}

                                                                             \subsubsection{Preparing to Compile}
                                                                             First enter our new directory:
                                                                             \begin{lstlisting}
                                                                             $ cd modules-3.2.10
                                                                             \end{lstlisting}
                                                                             Then read the \textit{README} and \textit{INSTALL} files, and, once you're done, read them once more just to 
                                                                             be safe. Once you are ready, I would advise you to run the following command to see how far it gets, the 
                                                                             rationale behind this is that, ultimately, every installation is different and all are missing different 
                                                                             things\footnote{In general, base Linux installations come with TCL but not TCLX and, thusly, it makes sense 
                                                                                 to not look for it, and, in any case, it will not be required by any of the scripts we write.}, therefore, 
                                                                                 it would make sense for you to see what the program is expecting in order to be compiled (alternatively you 
                                                                                         may skip this step entirely and pick up near our next command):
                                                                                     \begin{lstlisting}
                                                                                     $ ./configure --without-tclx
                                                                                     \end{lstlisting}
                                                                                     If it fails, then it means it is unable to find something that it needs in your system, and, well, this is a 
                                                                                     prime opportunity to get acquainted with the day to day life of a System Administrator. As such, the onus is 
                                                                                     on you to fix and procure anything and everything you may be missing for your installation. If, however, it 
                                                                                     finishes correctly, stop and don't do anything, for we haven't correctly configured it and it would be a 
                                                                                     mistake to install it as it is. What we must do, now, is run our configure script so that it set all the 
                                                                                     variables correctly and so our compilation process is as painless and easy as humanly possible.

                                                                                     Our next command is a bit more involved but by no means anything out of the ordinary (especially if you did 
                                                                                             in fact read both documents):
                                                                                         \begin{lstlisting}
                                                                                         $ ./configure --prefix=/share/apps/modules/3.2.10/\ 
                                                                                         --with-module-path=/share/modules/ --without-tclx\
                                                                                                            --enable-shell-funcs --enable-shell-alias --enable-shell-eval 
                                                                                                            \end{lstlisting}
                                                                                                            Once it's done you will now be ready to begin compiling\footnote{The more observant among you will note 
                                                                                                                that we will in fact be using the system's default compiler, and that it isn't (in general) the best idea.
                                                                                                                    However, keep in mind that in order to correctly manage several gcc versions we will need modules installed
                                                                                                                    and functioning, and, therefore, it makes absolute sense to simply set up modules with the base compiler. 
                                                                                                                    Additionally, we will not need anything more than a basic ANSI C compiler, and, thusly, our system's gcc
                                                                                                                    will suffice. Performance-wise, the same applies, the program is so simple and small that it is a true
                                                                                                                    fool's errand to pursue compilation of modules with any other compiler.} modules.

                                                                                                                    \subsubsection{Compiling Modules}
                                                                                                                    There isn't a lot to say about the actual compilation process, as it involves a very simple command that is 
                                                                                                                    in no way difficult to use:
                                                                                                                    \begin{lstlisting}
                                                                                                                    $ make
                                                                                                                    \end{lstlisting}
                                                                                                                    Once you enter the command, I'd advise a short walk or coffee break as it might take a short while. If it 
                                                                                                                    completes correctly the you are now ready to test your binary:
                                                                                                                    \begin{lstlisting}
                                                                                                                    $ make check
                                                                                                                    \end{lstlisting}
                                                                                                                    Now, if either the checking process, or the compilation process fails, it probably means you did something 
                                                                                                                    wrong and you'd be best served by returning to the beginning of this process and trying again. In theory, it 
                                                                                                                    shouldn't fail in any step of the way once the configure script passes, but nothing is absolutely assured to 
                                                                                                                    work.

                                                                                                                    \subsubsection{Installing Modules}
                                                                                                                    Having hurdled this enormous barrier, it is now time to finally install the products of our labor. In order 
                                                                                                                    to do this correctly I would advise you doing the following, as this will temporarily give you write access and permissions over the
                                                                                                                    folders where you will install the modules package and it will then return both the user and the group to root so that it can't be
                                                                                                                    modified by accident.
                                                                                                                    \begin{lstlisting}
                                                                                                                    $ sudo mkdir -p /share/apps/modules/3.2.10
                                                                                                                    $ sudo chown -R pi:pi /share/apps/modules
                                                                                                                    $ make install
                                                                                                                    $ sudo chown -R root:root /share/apps/modules
                                                                                                                    \end{lstlisting}

                                                                                                                    \subsubsection{Configuring modules}
                                                                                                                    Assuming the installation process has gone smoothly, the next step is quite simple, but necessary in order
                                                                                                                    to guarantee that modules will be available to each and every user of the cluster, no matter what machine
                                                                                                                    they use. With this in mind, first head to the folder where you installed modules, which, in our case, 
                                                                                                                    would be \textit{/share/apps/modules/3.2.10/}.
                                                                                                                    \begin{lstlisting}
                                                                                                                    $ cd /share/apps/modules/3.2.10
                                                                                                                    \end{lstlisting}
                                                                                                                    Once there, move to the folder where the initialization scripts are located, and verify that they indeed are
                                                                                                                    there.
                                                                                                                    \begin{lstlisting}
                                                                                                                    $ cd 3.2.10/init
                                                                                                                    $ ls
                                                                                                                    \end{lstlisting}
                                                                                                                    There, a file for all supported shells is available, and ready to be installed. In order to finish the set
                                                                                                                    up, all that needs to be done is run the following command.
                                                                                                                    \begin{lstlisting}
                                                                                                                    $ sudo cp modules.sh /etc/profile.d/.
                                                                                                                    \end{lstlisting}
                                                                                                                    Once that command completes, your modules installation should be up and ready to use. Simply reboot your
                                                                                                                    machine and log back in. Once you've done so, run the following command:
                                                                                                                    \begin{lstlisting}
                                                                                                                    $ module
                                                                                                                    \end{lstlisting}
                                                                                                                    If usage instructions appear, the your modules installation is complete and you are now ready to proceed
                                                                                                                    with the installation of any other package you may desire.

                                                                                                                    \subsection{Obtaining a new Compiler}
                                                                                                                    Raspbian, based on Debian wheezy, comes by default with GCC version 4.6.3, which has been thoroughly tested 
                                                                                                                    and is quite stable, but, unfortunately, it lacks support for most of the newer features of the C++ 
                                                                                                                    programming language, and therefore, we will need to upgrade to a newer version in order to make full 
                                                                                                                    use of these features\footnote{Do keep in mind that if you are using the Debian Jessie release, you will 
                                                                                                                        most likely have a newer compiler that does not suffer this issue.}. The version that has been chosen is
                                                                                                                        version 4.8.5, which is quite stable and supports all the features we need for the software we will be
                                                                                                                        installing in the future, additionally, it is very well supported and is far easier to install and use then
                                                                                                                        newer gcc releases, that have been known to be somewhat problematic on embedded systems.

                                                                                                                        \subsubsection{Overview of what will be done}
                                                                                                                        While it is true that precompiled packages do exist in the wild and that they are completely usable, 
                                                                                                                        it is a worthwhile exercise to compile your own compiler for your system, mostly due to the fact that 
                                                                                                                        going forward, in order to have the newest version and the latest  and greatest features, it is necessary 
                                                                                                                        to manually build it. Compiling our new compiler will be a somewhat more difficult task than compiling
                                                                                                                        Modules, simply because of the fact that there are a great deal of options and nuanced steps to be aware of.
                                                                                                                        However, despite all of that, in the end it will turn out to be a rather simple project, if all indeed does
                                                                                                                        according to plan.
                                                                                                                        \\
                                                                                                                            Do keep in mind that there are no real shortcuts to be taken in the compilation of gcc, but there several 
                                                                                                                            somewhat well defined steps to follow, which, in our case, will not suppose too great a challenge. The steps
                                                                                                                            that we shall follow are:
                                                                                                                            \begin{itemize}
                                                                                                                                \item Download and install the GNU Multiple Precision Arithmetic Library (GMP) library, version 6.1.0
                                                                                                                                    \item Download and install the GNU MPFR library, version 3.1.3
                                                                                                                                        \item Download and install the GNU MPC library, version 1.0.3
                                                                                                                                            \item Download and install the GNU Compiler Collection (GCC), version 4.8.5
                                                                                                                                            \end{itemize}

                                                                                                                                            \subsubsection{Obtaining the Software}
                                                                                                                                            We'll need to download the source from the official website, which, in it of itself is a quite simple task.
                                                                                                                                            To do this, all that needs to be done is to input the following commands:
                                                                                                                                            \begin{lstlisting}
                                                                                                                                            $ cd
                                                                                                                                            $ wget https://gmplib.org/download/gmp/gmp-6.1.0.tar.lz
                                                                                                                                            $ wget http://www.mpfr.org/mpfr-current/mpfr-3.1.3.tar.gz
                                                                                                                                            $ wget ftp://ftp.gnu.org/gnu/mpc/mpc-1.0.3.tar.gz
                                                                                                                                            $ wget http://mirrors-usa.go-parts.com/gcc/releases/\
                                                                                                                                                gcc-4.8.5/gcc-4.8.5.tar.gz
                                                                                                                                                $ tar --lzip -xvf gmp-6.1.0.tar.lz
                                                                                                                                                $ tar xzvf mpfr-3.1.3.tar.gz
                                                                                                                                                $ tar xzvf mpc-1.0.3.tar.gz
                                                                                                                                                $ tar xzvf gcc-4.8.5.tar.gz
                                                                                                                                                \end{lstlisting}
                                                                                                                                                Do make sure that you have successfully downloaded and extracted the correct compiler. If you so desire
                                                                                                                                                you may verify the integrity and validity of your download.

                                                                                                                                                \subsubsection{Preparing to Compile}
                                                                                                                                                As with any other software, compilation of gcc can be made as complex or as simple as is desired, though,
                                                                                                                                                usually, in order to make it work as well as can be in a given system, tuning of its compilation parameters
                                                                                                                                                is not only recommended, but actually, necessary. With this in mind, several resources must be read, at 
                                                                                                                                                least cursorily, in order to guarantee a smooth build process. Out of these resources, the one I would perhaps recommend reading through
                                                                                                                                                the most is, without a doubt, the GCC compiler's homepage, where, in every single corner, lies some useful gem of information. \\

                                                                                                                                                Fundamentally, compilers are complex and very demanding programs, that are more then capable of using every single bit of
                                                                                                                                                their host operating system, and, as such, are absolutely non-trivial tests of a system's setup and capabilities. In addition, they are
                                                                                                                                                more than likely to be difficult to get working correctly on a system, thus necessitating a rather involved set of steps to get them
                                                                                                                                                working. It is thus with this that we must now focus our attention on  actually beginning to compile our very own compiler. In order to
                                                                                                                                                set this process into motion we will need to compile and install several libraries so that we can even begin to build gcc on our system.

                                                                                                                                                \subsubsection{Compiling and installing GMP}
                                                                                                                                                The first of the aforementioned libraries is GMP (or GNU Multiple Precision Arithmetic Library). Fortunately for us, the build process 
                                                                                                                                                for it is rather painless, and, of course, we in theory already have made ourselves with all the requirements to get it running. Go 
                                                                                                                                                ahead and move the folder where it is located:
                                                                                                                                                \begin{lstlisting}
                                                                                                                                                $ cd gmp-6.1.0
                                                                                                                                                \end{lstlisting}
                                                                                                                                                Now that you've made it here, you will need to configure the program so that we can compile it, thus, let us do exactly like we did with
                                                                                                                                                modules and let's make sure that our configure script is capable of finishing with no options:
                                                                                                                                                \begin{lstlisting}
                                                                                                                                                $ ./configure
                                                                                                                                                \end{lstlisting}
                                                                                                                                                If it successfully finishes, then that means we are well on our way and that things have been correctly setup. We shall now run
                                                                                                                                                the configuration script once again, but, this time around, with the options we will actually use:
                                                                                                                                                \begin{lstlisting}
                                                                                                                                                $ ./configure --prefix=/share/apps/gmp/6.1.0/
                                                                                                                                                \end{lstlisting}
                                                                                                                                                We will be installing it following our convention, as it would otherwise not install it in the shared file system. In any case, it 
                                                                                                                                                should, once again, finish without a problem. Now that you've configured it, we must build the library itself:
                                                                                                                                                \begin{lstlisting}
                                                                                                                                                $ make
                                                                                                                                                \end{lstlisting}
                                                                                                                                                If it has finished correctly, and it appears to not have had any errors, then you are now ready to install it to our preferred location
                                                                                                                                                on our shared file system. To do so, enter the following commands:
                                                                                                                                                \begin{lstlisting}
                                                                                                                                                $ sudo mkdir -p /share/apps/gmp/6.1.0
                                                                                                                                                $ sudo chown -R pi:pi /share/apps/gmp
                                                                                                                                                $ make install
                                                                                                                                                $ sudo chown -R root:root /share/apps/gmp
                                                                                                                                                \end{lstlisting}
                                                                                                                                                As you can quite clearly see, this is very similar to what we did with modules.

                                                                                                                                                \subsubsection{Compiling and installing MPFR}
                                                                                                                                                Just as we had down before with GMP, we need to compile and install mpfr, which, fortunately for us, is done through pretty much the
                                                                                                                                                same process. Now, it is time to head into its folder:
                                                                                                                                                \begin{lstlisting}
                                                                                                                                                $ cd; cd mpfr-3.1.3
                                                                                                                                                \end{lstlisting}
                                                                                                                                                And, just as we had done before, we need to configure it, however, this library requires that GMP be present and compiled, so it is,
                                                                                                                                                therefore, advisable, to not just run \textit{configure} with nothing else added. In addition to that, we will include, at once where
                                                                                                                                                we will be installing it, in order to avoid redoing this step. With that in mind, run the following command:
                                                                                                                                                \begin{lstlisting}
                                                                                                                                                $ ./configure --with-gmp=/share/apps/gmp/6.1.0/ \ 
                                                                                                                                                --prefix=/share/apps/mpfr/3.1.3/
                                                                                                                                                \end{lstlisting}
                                                                                                                                                If you managed to do this without a problem, then those are more than phenomenal, as you've also now verified that you gmp installation 
                                                                                                                                                is correct and it works. If not, then you will most likely find answers as to why in, as always, the \textit{config.log} that you'll
                                                                                                                                                find in this very folder. Assuming that it did work correctly, run the following commands\footnote{I'll now skip over the lengthy 
                                                                                                                                                    explanation, as it is pretty much exactly the same as with GMP.}:
                                                                                                                                                    \begin{lstlisting}
                                                                                                                                                    $ make
                                                                                                                                                    $ sudo mkdir -p /share/apps/mpfr/3.1.3
                                                                                                                                                    $ sudo chown -R pi:pi /share/apps/mpfr
                                                                                                                                                    $ make install
                                                                                                                                                    $ sudo chown -R root:root /share/apps/mpfr
                                                                                                                                                    \end{lstlisting}
                                                                                                                                                    After you've successfully completed these steps, you must now move on to the next library needed for GCC.

                                                                                                                                                    \subsubsection{Compiling and installing MPC}
                                                                                                                                                    Like a broken record, we'll now perform a very similar procedure for MPC, the, thankfully, final library that will be needed in order to 
                                                                                                                                                    compile GCC. This time around the command to configure is just a tad longer, as it requires both GMP and MPFR to be compiled. Head into
                                                                                                                                                    the folder:
                                                                                                                                                    \begin{lstlisting}
                                                                                                                                                    $ cd; cd mpc-1.0.3
                                                                                                                                                    \end{lstlisting}
                                                                                                                                                    Now run the following configuration command:
                                                                                                                                                    \begin{lstlisting}
                                                                                                                                                    $ ./configure --with-gmp=/share/apps/gmp/6.1.0/ \
                                                                                                                                                                             --with-mpfr=/share/apps/mpfr/3.1.3/ --prefix=/share/apps/mpc/1.0.3
                                                                                                                                                                             \end{lstlisting}
                                                                                                                                                                             If it completes you are now ready to install it, which follows more or less the same pattern as before\footnote{Interestingly, this
                                                                                                                                                                                 isn't in any way a bad thing, as one of the principles of the \textit{GNU autoconf}/\textit{automake} system is homogeneity across the
                                                                                                                                                                                     programs deployed with it, in the sense that they should all follow a similar installation procedure and be reasonably simple to
                                                                                                                                                                                     get into a working state.}, and will require you to input these commands:
                                                                                                                                                                                     \begin{lstlisting}
                                                                                                                                                                                     $ make
                                                                                                                                                                                     $ sudo mkdir -p /share/apps/mpc/1.0.3
                                                                                                                                                                                     $ sudo chown -R pi:pi /share/apps/mpc
                                                                                                                                                                                     $ make install
                                                                                                                                                                                     $ sudo chown -R root:root /share/apps/mpc
                                                                                                                                                                                     \end{lstlisting}
                                                                                                                                                                                     Fortunately, if everything up to this point has been successfully completed, then, our goal of installing gcc is quite close, however,
                                                                                                                                                                                     there is still some work to be done.

                                                                                                                                                                                     \subsubsection{Compiling and installing GCC}
                                                                                                                                                                                     At last we've come to face the great challenge that awaits us, compiling the behemoth that is GCC. Though, now being armed with all the
                                                                                                                                                                                     libraries we could possibly need, we are now completely ready to embark on this long journey. At first, it'll all be quite simple, and 
                                                                                                                                                                                     eerily similar to what we've faced before with all of the libraries. However, this time things will be a bit more involved. Now, in
                                                                                                                                                                                     order to begin first head to where gcc is:
                                                                                                                                                                                     \begin{lstlisting}
                                                                                                                                                                                     $ cd; cd gcc-4.8.5
                                                                                                                                                                                     \end{lstlisting}
                                                                                                                                                                                     Once you are here, you must first create a new folder and move inside it:
                                                                                                                                                                                     \begin{lstlisting}
                                                                                                                                                                                     $ mkdir objdir
                                                                                                                                                                                     $ cd objdir
                                                                                                                                                                                     \end{lstlisting}
                                                                                                                                                                                     This is necessary as GCC cannot be built right in its main directory, and as such, this is a convenient and simple solution to this
                                                                                                                                                                                     issue. Now, it is time to begin configuring GCC, and, before we do, I need to warn you of the following: \textbf{GCC, being such a
                                                                                                                                                                                         complex application, is quite difficult to configure and compile correctly on the first try, and it took me quite some time to find
                                                                                                                                                                                             which options undoubtedly worked and which didn't with my particular raspberry pi system. As such, do not be alarmed if the 
                                                                                                                                                                                             instructions you find here do not work perfectly on your system, as it is perfectly possible that some minute detail has been 
                                                                                                                                                                                             overlooked on either of our ends, or you have tweaked your system in some way that doesn't allow you to build with these specific
                                                                                                                                                                                             instructions.} With this out of the way, let us begin, run the following command, which, in theory, already has been tailored to work
                                                                                                                                                                                             well with the Raspberry Pi:
                                                                                                                                                                                             \begin{lstlisting}
                                                                                                                                                                                             $ ../configure --with-gmp=/share/apps/gmp/6.1.0/ \
                                                                                                                                                                                                                       --with-mpfr=/share/apps/mpfr/3.1.3/ \
                                                                                                                                                                                                                       --with-mpc=/share/apps/mpc/1.0.3 \
                                                                                                                                                                                                                       --enable-languages=c,c++,fortran --disable-multilib \
                                                                                                                                                                                                                       --enable-shared --enable-__cxa_atexit \
                                                                                                                                                                                                                       --enable-threads=posix --enable-nls \
                                                                                                                                                                                                                       --enable-gnu-unique-object --with-arch=armv6 \
                                                                                                                                                                                                                       --with-fpu=vfp --with-float=hard \
                                                                                                                                                                                                                       --prefix=/share/apps/gcc/4.8.5/
                                                                                                                                                                                                                       \end{lstlisting}
                                                                                                                                                                                                                       I don't exactly recall how long this should take, but be sure that it won't be a quick process, therefore, I would advise that you turn
                                                                                                                                                                                                                       your attention to some other task until you see this finish, and don't forget to be patient. If it fails, you'll probably
                                                                                                                                                                                                                       find the answer as to what happened in the \textit{config.log} file that's left afterwards. If it doesn't, then it means that you are 
                                                                                                                                                                                                                       well on your way to compiling GCC. The next step to be taken is quite simple, though it is arguably the most difficult:
                                                                                                                                                                                                                       \begin{lstlisting}
                                                                                                                                                                                                                       $ make
                                                                                                                                                                                                                       \end{lstlisting}
                                                                                                                                                                                                                       While at first glance it seems to make no sense that this step is difficult, it must be understood that the make process might fail and
                                                                                                                                                                                                                       it might be quite difficult to recover from those errors, nonetheless, it is absolutely solvable, and, given enough time, I can assure 
                                                                                                                                                                                                                       you that you will find an adequate solution to the issue. Now, another thing that must be said is that you will need to be very patient,
                                                                                                                                                                                                                       more than you could probably imagine, so don't despair if it seems as if it's taking forever. Once it's done, there's quite a great
                                                                                                                                                                                                                       many things to be done. The very next thing you need to do is quite simple, and, as ever, quite similar to what we've done:
                                                                                                                                                                                                                       \begin{lstlisting}
                                                                                                                                                                                                                       $ sudo mkdir -p /share/apps/gcc/4.8.5
                                                                                                                                                                                                                       $ sudo chown -R pi:pi /share/apps/gcc
                                                                                                                                                                                                                       $ make install
                                                                                                                                                                                                                       $ sudo chown -R root:root /share/apps/gcc
                                                                                                                                                                                                                       \end{lstlisting}
                                                                                                                                                                                                                       Once again, be patient while the application installs, as it might take a while. If you did these steps correctly, you will now be able
                                                                                                                                                                                                                       to do a simple tests to see if things are, at least in some way, correctly installed and ready to use:
                                                                                                                                                                                                                       \begin{lstlisting}
                                                                                                                                                                                                                       $ cd /share/apps/gcc/4.8.5/bin
                                                                                                                                                                                                                       $ ./gcc -v
                                                                                                                                                                                                                       \end{lstlisting}
                                                                                                                                                                                                                       You should see some output stating the program's version and how it was configured, among other things. If you do, then you did things,
                                                                                                                                                                                                                       so far, correctly.

                                                                                                                                                                                                                       \subsubsection{Preparing a modulefile for GCC}
                                                                                                                                                                                                                       Now it is time to create our very own module for gcc, as, going forward, we will need to be able to select easily which version we want
                                                                                                                                                                                                                       without having to fiddle around with manually changing environment variables. As such, I will now leave you with the text to the module
                                                                                                                                                                                                                       and you would do well to use it, though please be sure to look up the manual to the modulefiles so you understand what comes next:
                                                                                                                                                                                                                       \begin{lstlisting}
#%Module1.0#################################################
##
## Modulefile for Gcc 4.8.5
##
##
                                                                                                                                                                                                                       proc ModulesHelp {} {
                                                                                                                                                                                                                            global version modroot
                                                                                                                                                                                                                                    
                                                                                                                                                                                                                                    puts stderr "Prepares the environment to use Gcc 4.8.5,
                                                                                                                                                                                                                                                        and loads all of it's libraries aswell"
                                                                                                                                                                                                                       }

module-whatis "Prepares the environment to use GCC 4.8.5"

set     topdir          /share/apps/gcc/4.8.5
set     version         4.8.5
set     sys         armv6

module load gmp/6.1.0 mpfr/3.1.3 mpc/1.0.2

setenv      CC          $topdir/bin/gcc
setenv      GCC         $topdir/bin/gcc
setenv      FC          $topdir/bin/gfortran
setenv      F77         $topdir/bin/gfortran
setenv      F90         $topdir/bin/gfortran

prepend-path    C_INCLUDE_PATH      $topdir/include
prepend-path    CPLUS_INCLUDE_PATH  $topdir/include
prepend-path    PATH            $topdir/bin
prepend-path    MANPATH         $topdir/share/man
prepend-path    LD_LIBRARY_PATH $topdir/lib
prepend-path    LD_RUN_PATH     $topdir/lib
prepend-path    LIBRARY_PATH    $topdir/lib
###########################################################
\end{lstlisting}
This, as it stands, isn't ready to be used yet since we need to create the modules for GMP, MPFR and MPC, and only then would the new
GCC be truly ready to be used. The first step you must take is to navigate to the following folder:
\begin{lstlisting}
$ cd /share/apps/modules/3.2.10/modulefiles
\end{lstlisting}
Now, you must create a folder for your new apps modulefile, and create a blank file which you will now edit:
\begin{lstlisting}
$ sudo mkdir gcc
$ cd gcc
$ sudo touch 4.8.5
\end{lstlisting}
Now that it's ready, it's time to add the module's text to the file:
\begin{lstlisting}
$ sudo chown pi:pi 4.8.5
$ vi 4.8.5
$ sudo chown root:root 4.8.5
\end{lstlisting}
After you've done all of this, you are now ready to proceed to creating all the modulefiles that you need to do next.

\subsubsection{Preparing a modulefile for GMP, MPFR and MPC}
In order to finally get GCC working, and allowing you to use it's module, we need to create modules for our GMP, MPFR, and MPC
libraries. As such, go ahead and create folders for all of the modules:
\begin{lstlisting}
$ cd ..
$ sudo mkdir {gmp,mpfr,mpc}
$ sudo touch {gmp/6.1.0,mpfr/3.1.3,mpc/1.0.3}
$ sudo chown -R pi:pi {gmp,mpfr,mpc}
$ vi gmp/6.1.0
\end{lstlisting}
At this point, insert the following text into the \textit{6.1.0} file:
\begin{lstlisting}
#%Module1.0#################################################
##
## Modulefile for GMP 6.1.0
##
##
proc ModulesHelp {} {
        global version modroot
                
                puts stderr "Prepares the environment to use GMP 6.1.0"
}

module-whatis "Prepares the environment to use GMP 6.1.0"

set     topdir          /share/apps/gmp/6.1.0
set     version         6.1.0
set     sys         armv6

prepend-path    C_INCLUDE_PATH      $topdir/include
prepend-path    CPLUS_INCLUDE_PATH  $topdir/include
prepend-path    LD_LIBRARY_PATH     $topdir/lib
prepend-path    LD_RUN_PATH     $topdir/lib
prepend-path    LIBRARY_PATH        $topdir/lib
###########################################################
\end{lstlisting}
Once you've done this, move on to the next file:
\begin{lstlisting}
$ vi mpfr/3.1.3
\end{lstlisting}
At this point, insert the following text into the \textit{3.1.3} file:
\begin{lstlisting}
#%Module1.0#################################################
##
## Modulefile for MPFR 3.1.3
##
##
proc ModulesHelp {} {
        global version modroot
                
                puts stderr "Prepares the environment to use MPFR 3.1.3"
}

module-whatis "Prepares the environment to use MPFR 3.1.3"

set     topdir          /share/apps/mpfr/3.1.3
set     version         3.1.3
set     sys         armv6

prepend-path    C_INCLUDE_PATH      $topdir/include
prepend-path    CPLUS_INCLUDE_PATH  $topdir/include
prepend-path    LD_LIBRARY_PATH     $topdir/lib
prepend-path    LD_RUN_PATH     $topdir/lib
prepend-path    LIBRARY_PATH        $topdir/lib
###########################################################
\end{lstlisting}
Once you've done this, move on to the next file:
\begin{lstlisting}
$ vi mpc/1.0.3
\end{lstlisting}
At this point, insert the following text into the \textit{1.0.3} file:
\begin{lstlisting}
#%Module1.0#################################################
##
## Modulefile for MPC 1.0.3
##
##
proc ModulesHelp {} {
        global version modroot
                
                puts stderr "Prepares the environment to use MPC 1.0.3"
}

module-whatis "Prepares the environment to use MPC 1.0.3"

set     topdir          /share/apps/mpc/1.0.3
set     version         1.0.3
set     sys         armv6

prepend-path    C_INCLUDE_PATH      $topdir/include
prepend-path    CPLUS_INCLUDE_PATH  $topdir/include
prepend-path    LD_LIBRARY_PATH     $topdir/lib
prepend-path    LD_RUN_PATH     $topdir/lib
prepend-path    LIBRARY_PATH        $topdir/lib
###########################################################
\end{lstlisting}
Now, if you've done all these steps correctly, you will be able to see new items pop up when you run the following command:
\begin{lstlisting}
$ module avail
\end{lstlisting}
It should show you that all of our new modules are now ready to be used. Go ahead and test this by doing the following:
\begin{lstlisting}
$ module load gcc/4.8.5
$ module list
\end{lstlisting}
You should now see that the GCC module is loaded along with the GMP, MPFR and MPC modules. Finally, test that everything went right with
the following command:
\begin{lstlisting}
$ gcc -v
\end{lstlisting}
If all went well, then you should see it say that you are running GCC version 4.8.5, and no errors should show up. And with that, you'd
be done with the build and installation of GCC. If you are so inclined, and depending on how much space your SD card has, I would advise
you to delete the sources and extracted folders \textbf{(but not the installed folders in \textit{/share}!)} for GCC, it's libraries and 
for Modules, though this is entirely up to you.

\subsection{Obtaining an MPI implementation}
Our machine is now equipped to run and compile C and C++ sources that require some of the newer features of these languages.
Unfortunately we're not quite yet ready to begin preparing and deploying parallel applications, as we have no way for the applications
to talk to one another whilst they're running on different machines. This is where MPI comes in, as it defines a message passing
interface that standardizes how programs ought to talk to each of their instances when they're running on parallel machines. In terms
of what this means for us, well, it means we will need another layer of software on top of our compiler, and its associated libraries 
and headers in order to compile the parallel programs we will want to run.

\subsubsection{Overview of what will be done}
In a stroke of luck for us, setting up our MPI implementation isn't truly that difficult, and it, thankfully, relies on the now familiar
paradigm of the \textit{automake}/\textit{autoconf} system. In this case, we've chose to stick to a tried and true MPI implementation,
         which is OpenMPI version 1.10.2, which is the latest stable release. As with what we've done before, we'll delve into the world of
         compilation once more in order to build this software just how we want it. With this in mind, what we'll do is the following:
         \begin{itemize}
         \item We'll download and obtain the source for OpenMPI version 1.10.2
         \item We'll compile it and configure it to our liking
         \item We'll install it under \textit{/share} so that in due time, all of our nodes can access it.
         \end{itemize}

         \subsubsection{Obtaining the Software}
         In order to get the latest version of OpenMPI, we'll need to download it from their website with the following set of commands (and
                 we'll once again download everything to our home):
             \begin{lstlisting}
             $ cd
             $ wget https://www.open-mpi.org/software/ompi/v1.10/\
                 downloads/openmpi-1.10.2.tar.gz
                 $ tar xzvf openmpi-1.10.2.tar.gz
                 \end{lstlisting}
                 Now that you've extracted it, it is time to prepare to compile it.

                 \subsubsection{Preparing to compile}
                 Interestingly, there really aren't a great deal of new things to take into consideration once we begin the journey of building OpenMPI,
                 and as such, it makes sense to say that pretty much all of the warnings and considerations that applied to GCC apply to OpenMPI, though,
                 in this case, with a minimally customized configuration and our newly installed GCC things should compile rather smoothly. If, for some
                 reason OpenMPI has strange configuration issues, you'd do yourself a great favor, as always, by checking your \textit{config.log}. With
                 not much else, then to say, we are ready to begin.

                 \subsubsection{Compiling and installing OpenMPI}
                 First, head into the folder you've extracted:
                 \begin{lstlisting}
                 $ cd; cd openmpi-1.10.2
                 \end{lstlisting}
                 Once you're there, I'd advise reading through the \textit{README}, for it may contain interesting information that may be directly 
                 relevant to you particular use of the cluster. Once you're done with that you may realize that it is then time to begin configuring the
                 software to our specific system's needs. This time, not too may options will be needed to be used to produce a fully functional 
                 OpenMPI build. I must, nonetheless, advise that as time goes on you come to find a version that you trust and find completely stable
                 on your particular system, as, in due time, you'll find that it is nearly inveitable to find bugs in the OpenMPI you have installed, 
                 and, this willingness to change is quite important and useful, even if it is just in search of performance enhancements\footnote{This, 
                     though, does not mean that the version we'll use is in any way unsable or bad.}. Thus if you've come to terms with this, I may suggest 
                     that you do the following\footnote{This time around you'll probably notice that I'll use a different and quite strange foldering 
                         convention for our new OpenMPI, and, the reason for this is rather simple, as time goes on and the needs of your cluster change, you may 
                             be inclined to test the same version of openmpi with a different compiler, and, as such, it becomes important to be to easily file the 
                             software as you go along. This practice is inspired by how our University's cluster has done things, and it has worked wonders, as it 
                             greatly simplifies searching for the specific piece of software we are looking for, though, of course, you are free to change this up 
                             however you want. It also somewhat eases scripting, if you ever need to do such things.}:
                             \begin{lstlisting}
                             $ ./configure --prefix=/share/apps/openmpi/1.10.2/gcc/4.8.5/\
                                                    --build=arm-linux-gnueabihf
                                                    \end{lstlisting}
                                                    This configuration will probably take some time, but, in theory should not fail, provided you've loaded the correct modules and made
                                                    sure that you inputted all of the parameters correctly. The next step is somewhat different to what we've done before, as it is
                                                    recommended that you do both the build and installation in one step, so we'll prepare our folders before we do that and then begin
                                                    compiling:
                                                    \begin{lstlisting}
                                                    $ sudo mkdir -p /share/apps/openmpi/1.10.2/gcc/4.8.5
                                                    $ sudo chown -R pi:pi /share/apps/openmpi
                                                    $ make all install
                                                    $ sudo chown -R root:root /share/apps/openmpi
                                                    \end{lstlisting}
                                                    If all of these steps complete without an issue, then the software has been installed. Verify this by running the following command and
                                                    checking that the output is indeed of the GCC you compiled it with:
                                                    \begin{lstlisting}
                                                    $ /share/apps/openmpi/1.10.2/gcc/4.8.5/bin/mpicc -v
                                                    \end{lstlisting}
                                                    Once you've verified that all is well, it is now time to realize that work still has to be done before we can use our new OpenMPI.As you
                                                    might imagine, we'll create a new module for it, that will allow us to use it with ease.

                                                    \subsubsection{Preparing a modulefile for OpenMPI}
                                                    Just as we had to do with GCC, we will need to create a module for OpenMPI, so that we may simplify our lives just a bit whenever we are 
                                                    to use it. With that in mind, let us do the following\footnote{Once again you'll notice a difference in how the module is named, as it
                                                        is quite important to know exactly what a module really is loading, and it isn't an issue as modules does have autocomplete built in}:
                                                        \begin{lstlisting}
                                                        $ cd /share/apps/modules/3.2.10/modulefiles
                                                        $ sudo mkdir openmpi
                                                        $ sudo touch openmpi/1.10.2_gcc-4.8.5
                                                        $ sudo chown pi:pi openmpi/1.10.2_gcc-4.8.5
                                                        $ vi openmpi/1.10.2_gcc-4.8.5
                                                        \end{lstlisting}
                                                        This new file we'll be constructing will be somewhat longer but in now way really different to what we're used to, so with that in mind, 
                                                        input the following into your file:
                                                        \begin{lstlisting}
#%Module1.0#################################################
##
## Modulefile for OpenMPI 1.10.2 built with GCC 4.8.5
##
##
                                                        proc ModulesHelp {} {
                                                                global version modroot
                                                                        
                                                                        puts stderr "Prepares the environment to use OpenMPI 1.10.2"
                                                        }

module-whatis "Prepares the environment to use OpenMPI 1.10.2"

set     topdir  /share/apps/openmpi/1.10.2/gcc/4.8.5
set     version         1.10.2
set     sys         armv6

module load gcc/4.8.5

prepend-path    PATH            $topdir/bin
prepend-path    C_INCLUDE_PATH      $topdir/include
prepend-path    CXX_INCLUDE_PATH    $topdir/include
prepend-path    CPLUS_INCLUDE_PATH  $topdir/include
prepend-path    LD_LIBRARY_PATH     $topdir/lib
prepend-path    LD_RUN_PATH     $topdir/lib
prepend-path    LIBRARY_PATH        $topdir/lib
prepend-path    MANPATH         $topdir/share/man
setenv      MPI_BIN         $topdir/bin
setenv      MPI_SYSCONFIG       $topdir/etc
setenv      MPI_RUN         $topdir/bin
setenv      MPI_INCLUDE     $topdir/include
setenv      MPI_LIB         $topdir/lib
setenv      MPI_MAN         $topdir/share/man
setenv      MPI_COMPILER        mpicc
setenv      MPI_SUFFIX          _openmpi
setenv      MPI_HOME            $topdir
setenv      CC                  mpicc
setenv      CXX                 mpiCC
setenv      F77                 mpifort
setenv      F90                 mpifort
setenv      FC                  mpifort
setenv      MPI_INTERCONNECT    p4
setenv      MPI_VENDOR      openmpi
###########################################################
\end{lstlisting}
As you can see, this particular modulefile is somewhat long, but is nonetheless very useful, as it contains most environment variables
you might need at some point with OpenMPI. Going forward we will probably make quite a bit of use of OpenMPI, so these steps have now
saved us quite a bit of time.

\subsection{Configuring your network}
Up to this point your raspberry pi has, most likely, been using a dynamic ip address, and this, unfortunately is not how things should
be, as it is quite useful to actually know where your node is on the network\footnote{Now this brings up a very interesting and 
    difficult thing about creating a cluster with Raspberry Pis. They typically only have one standard network interface to work with, 
              however, in a cluster, you should ideally have two networks, one that is connected to the outside world and isn't. The basic idea goes
                  something like this: your master node should be connected to the internet as it will most likely need it at some point, in fact, we
                  have used it quite a bit; however, you compute nodes should not be connected to the internet, as they can and will receive everything 
                  they need from your master node. Unfortunately this isn't really possible unless you have some sort of external network interface and
                  an additional network switch to create your secondary network. Given these limitations, it should then come as no surprise that we
                  will connect all of our nodes to the same switch/router, and therefore they will also have access to the internet, fortunately, this
                  really is no problem for us.}. With this in mind we will need to modify our network configuration and make some assumptions that may
                  be beneficial to us. For now, however, we will do a bare bones process that will allow us to prepare our cluster for it's initial
                  testing, and will, hopefully, give you the insight necessary to continue expanding the cluster later on.

                  \subsubsection{Setting up a static IP}
                  First of all, if possible, verify which IP's in your router's LAN have been taken, and select one that doesn't conflict with any of
                  those. In my case (and hopefully yours) we found \textit{192.128.1.130} to be a perfectly acceptable address to give to our master node. 
                  Therefore, we will now change our network configuration to make this our machine's default state:
                  \begin{lstlisting}
                  $ cd /etc/network/
                  $ sudo vi interfaces
                  \end{lstlisting}
                  Now, once you've opened up this file, modify it so that it looks like this:
                  \begin{lstlisting}
                  auto lo

                  iface lo inet loopback
                  iface eth0 inet static
                  address 192.168.1.130
                  netmask 255.255.255.0
                  gateway   192.168.1.1

                  allow-hotplug wlan0
                  iface wlan0 inet manual
                  wpa-roam /etc/wpa_supplicant/wpa_supplicant.conf
                  iface default inet dhcp
                  \end{lstlisting}

                  \subsubsection{Modifying your \textit{/etc/hosts} file}
                  Now, another important aspect of a cluster is the fact that these machines should be able to resolve a simple hostname in order to
                  simplify connecting and setting up MPI machine files, this, also, has the benefit of making it easier for us to know where exactly we
                  are connecting, and it also removes a large part of the guessing game that would come with just using IP addresses. To this end we must 
                  then modify our \textit{/etc/hosts} file\footnote{Interestingly, distributions such as Rocks have an automated \textit{hosts} 
                      modification system that automatically adds every new machine to every node's \textit{hosts} file, which greatly simplifies the 
                          administration of the cluster.}. Now, in order to do this, we must first decide where we will place our next nodes on our network, and 
                          how we will do our naming scheme. In this case, we will follow the standard convention of \textit{compute-X-Y}, where X indicates what 
                          rack it is on and Y indicates its position on the rack itself. Additionally, we will start assigning our new node's IP's from 
                          \textit{192.168.1.131} onwards. Therefore, at the very least we will modify our master node's \textit{/etc/hosts} file with at least one
                          node. In order to do this do the following:
                          \begin{lstlisting}
                          $ cd /etc/
                          $ sudo vi hosts
                          \end{lstlisting}
                          Once you've opened it, go ahead and add the following to the very end of the file:
                          \begin{lstlisting}
                          192.168.1.131 compute-0-0
                          \end{lstlisting}
                          This, now, will allow you to type out compute-0-0 (when we create it) so that you don't have to know each node's IP address by heart. As
                          you can now probably imagine, this will greatly simplify using ssh and filling out which computers you want to run parallel programs on.

                          \subsection{Final steps of the basic setup}
                          Now that you've finished doing all of these long and admittedly arduous tasks, you are now very nearly done. The next steps I'd 
                          recommend doing is some basic cleanup on all of the places you've worked, and some basic re-verification that all of the software is
                          working as it is intended and that the modules work correctly. One easy way to do this is to check what you environment variables look
                          like after you load the OpenMPI module, as they should, in theory, be full of the paths you've set for them in your scripts. You can
                          easily do this as follows:
                          \begin{lstlisting}
                          $ module load openmpi/1.10.2_gcc-4.8.5
                          $ env | less
                          \end{lstlisting}
                          Once you are satisfied with what you see there, and you are somewhat sure that all works as it should, I would then say it is now 
                          advisable to reboot your device so that the changes you've made to your network configuration are made permanent. You can easily do
                          this as follows:
                          \begin{lstlisting}
                          $ sudo reboot
                          \end{lstlisting}
                          After this you can now confidently state that you have functional (if a bit bare-bones) master node. Unfortunately we don't really have
                          any cool applications to run on it and test that seductively mysterious MPI functionality over our network. To fix this we must first
                          set up for ourselves a second node, that will act as our trusty computation node\footnote{As time passes by and, hopefully, your love 
                              and ambition for your own cluster grows, you will begin to think of a ton of ways you can employ new nodes. For example, you could have 
                                  one node as a sort of firewall for your master node, another as your storage node, a couple as dedicated computation nodes, another as 
                                  an account authentication node, and so on and so forth. Truly, the only limits to this are your imagination and your budget. If you're
                                  feeling specially adventurous you may want to even try out creating complex and novel network arrangements.}. After we've done that, 
                                  then we can finally go and set up some amazing things to see just how powerful and versatile our cluster is.

                                  \section{Setting up a secondary node}
                                  Now that our master node's basic setup is done, it is now time to begin preparing our first compute node, for one very  important 
                                  reason: we will need it to test our parallel applications, and it is therefore quite useful to have it ready. Fortunately it is
                                  not a very difficult process, as, differing from what we've done before, we will not need to a lot of things in order to prepare
                                  our node. Nodes, in theory, should not have a lot done to them aside from a very basic setup process, since, from now on, they will
                                  get all of the things they need from our shared storage. It is therefore advisable to not do too much to them in order to not clutter
                                  them with unnecessary things. With this in mind, we will begin with what we need to do.

                                  \subsection{Getting the latest versions of the default packages}
                                  Similarly to what we did on our master node, you should have a base and clean raspbian image mounted on this Raspberry Pi, there, once 
                                  you've done that, proceed to login using the same default user and password (user=\textit{pi}, password=\textit{raspberry}), and make 
                                  sure you are now successfully logged in. Now, run the following commands to update all of the base and default packages:
                                  \begin{lstlisting}
                                  $ sudo apt-get update
                                  $ sudo apt-get upgrade
                                  \end{lstlisting}
                                  Once that finishes, you are now ready to proceed to setting your node up.

                                  \subsection{Running the configuration script}
                                  This time around you, just as before, will need to run Raspbian's included configuration script so that ssh is enabled from now on.
                                  In addition to that, you will also need to change this Raspberry Pi's hostname, in order to avoid confusion and problems. To do this,
                                  do just as before and run the script:
                                  \begin{lstlisting}
                                  $ sudo raspi-config
                                  \end{lstlisting}
                                  Once you are there, under the advanced options, you will find the options you need to do the aformentioned steps. Go ahead and do that 
                                  and let the device reboot after you do so.

                                  \subsection{Installing all the needed utilities}
                                  For now, all we will truly need on this node is the ability to see the files that our master node has on its shared drive. For this we
                                  will need some utilities, though they are very easy to install, and will allow us to get our pi in working order in no time. To do this
                                  run the following command:
                                  \begin{lstlisting}
                                  $ sudo apt-get install rpcbind nfs-common
                                  \end{lstlisting}
                                  After it succesfully installs it, we will need to enable a service on our system, as, without it, we will be unable to mount our new
                                  filesystem, and, additionally we will want it to start as soon as we boot. With that in mind, run the following commands:
                                  \begin{lstlisting}
                                  $ sudo service rpcbind start
                                  $ sudo update-rc.d rpcbind enable
                                  \end{lstlisting}
                                  At this point we are now able to mount the NFS file system we set up at the very beginning on our master node.

                                  \subsection{Mounting the master node's NFS share}
                                  By now you've probably noticed that I really didn't install too many things on our new node and instead only sought to prepare it to 
                                  mount the master node's \textit{/share/} folder on our system. This is mainly due to the fact that it makes no sense to clutter up this
                                  node with unnecessary applications that will be shared anyways, and as such, it is a much smarter move to things this way. 
                                  Additionally, it allows one to learn about something that basically all modern clusters do. In order to this in a permanent way, 
                                  allowing us to always have it available when the device turns on, we will need to follow a couple of steps. First, create an empty 
                                  folder where you'll mount the remote share:
                                  \begin{lstlisting}
                                  $ cd /
                                  $ sudo mkdir share
                                  \end{lstlisting}
                                  With that done, go over to your \textit{/etc/} folder and edit your \textit{/etc/fstab} file:
                                  \begin{lstlisting}
                                  $ cd /etc/
                                  $ sudo vi fstab
                                  \end{lstlisting}
                                  At the end of the file add a line as follows:
                                  \begin{lstlisting}
                                  192.168.1.130:/share/ /share  nfs     0   0
                                  \end{lstlisting}
                                  This, from now on, will automatically mount the file system on boot, but, for now, we'll do it manually so we don't have to reboot the 
                                  device\footnote{Don't worry, this will in no way break what we've done with \textit{/etc/fstab}, it will simply allow us to begin 
                                      working from now on with our shared file system.}. So in order to do this, enter the following command:
                                      \begin{lstlisting}
                                      sudo mount -t nfs 192.168.1.130:/share /share
                                      \end{lstlisting}
                                      Now, go ahead and enter into the folder and verify that our apps folder is there and that you can now see all the apps we compiled and
                                      created back on our master node. You can do this as follows:
                                      \begin{lstlisting}
                                      $ cd /share
                                      $ ls
                                      \end{lstlisting}
                                      If you can both access and read the files you now mounted on \textit{/share}, then it all went well, and you are ready to proceed.

                                      \subsection{Preparing modules}
                                      Now that we can access the files on our master node, it is time to add some of the functionality that is on the main server onto our 
                                      new node, most notably, we might, at some point, require modules on it, so it's a great idea to prepare this computer to use them on 
                                      demand. Fortunately, now that we can access them remotely, we do not need to compile anything as it's been already done for us, this
                                      then, has the fantastic added benefit that we can now simply do some simple setup steps and work far quicker than before. To attain 
                                      this, we must run some simple commands:
                                      \begin{lstlisting}
                                      $ sudo cp /share/apps/modules/3.2.10/3.2.10/init/\
                                          modules.sh /etc/profile.d
                                          $ source /etc/profile.d/modules.sh
                                          \end{lstlisting}
                                          This will now allow us to do full use of modules whilst on our node, and we can immediately test this by doing the following:
                                          \begin{lstlisting}
                                          $ module avail
                                          \end{lstlisting}
                                          This command should now let us see all the modules that we now have at our disposal.

                                          \subsection{Configuring your network}
                                          The next, and extremely important step to be taken, is to correct our pi's network configuration, just as we had done with our master 
                                          node. The reason for this is simple, it is far simples to work with machines that have a statically assigned IP address, as it'll ease
                                          the work needed to both use and find them. As you might imagine, we'll need to modify our \textit{/etc/network/interfaces} file and 
                                          we'll need to modify our \textit{/etc/hosts}, though, just as before, this will not be complicated. Ideally as time goes on and we add 
                                          more devices, we'll simplify this process and hopefully automate it, but, for now, we'll do it manually.

                                          \subsubsection{Setting up a static IP}
                                          Just as we had done before, we'll need to do some modification to out \textit{/etc/network/interfaces} file, in the exact same manner 
                                          that we did with our master node. To do so, follow the following steps:
                                          \begin{lstlisting}
                                          $ cd /etc/network
                                          $ sudo vi interfaces
                                          \end{lstlisting}
                                          Edit the file so that it looks as follows:
                                          \begin{lstlisting}
                                          auto lo

                                          iface lo inet loopback
                                          iface eth0 inet static
                                          address 192.168.1.130
                                          netmask 255.255.255.0
                                          gateway   192.168.1.1

                                          allow-hotplug wlan0
                                          iface wlan0 inet manual
                                          wpa-roam /etc/wpa_supplicant/wpa_supplicant.conf
                                          iface default inet dhcp
                                          \end{lstlisting}
                                          Once you've done that, it is now time to take a look at your \textit{/etc/hosts} file.

                                          \subsubsection{Modifying your \textit{/etc/hosts} file}
                                          Running the risk of sounding like a broken record, it necessary to once again do this procedure, as it is simply a good practice, and
                                          it doesn't really take any effort. To do this move to your \textit{/etc} folder and open your \textit{hosts} file:
                                          \begin{lstlisting}
                                          $ cd /etc/
                                          $ sudo vi hosts
                                          \end{lstlisting}
                                          Next insert the following into the file at its end:
                                          \begin{lstlisting}
                                          192.168.1.130 raspberrypi
                                          \end{lstlisting}
                                          Once you've done this, the node is now, in theory, ready to be used.

                                          \subsection{Final steps of the basic setup}
                                          There is really not much to be done to the node now that you've gotten it up and running. Just make sure you don't have too much of its 
                                          local filesystem and make sure that you don't have any unnecessary application launching into the background. After you've done all 
                                          these checks, it is time to reboot it to apply your new network configuration and test your initialization configurations. So run:
                                          \begin{lstlisting}
                                          $ sudo reboot
                                          \end{lstlisting}
                                          Once it reboots you should see that the NFS share has been successfully mounted, and that you now have a static IP address. Once you 
                                          are happy with this, you may now say that you do indeed have two working nodes in a Raspbery Pi Cluster. From now on we will focus on 
                                          actually testing our cluster's capabilities, and making sure that it actually works as we intended. Though we might encounter some 
                                          issues along the way, I'm completely sure that none are too large in any single way to stop us now!

                                          \section{MPI, Stability and Performance Tests}
                                          Now that we have all of the basic functionalities done and ready to be used, a wonderful time comes, we are going to actually begin
                                          compiling and using powerful apps that will allow us to see just how well all our work went. This time around, we will focus on two 
                                          quite practical and interesting tests that will help us measure if our newly created infrastructure is actually any good. The first of
                                          these is a simple application whose only job is to calculate Pi\footnote{Which, in a way, is quite appropriate for this project} 
                                          utilizing parallel methods; meanwhile, the second focuses on giving an accurate reading of the peak performance of the cluster as a 
                                          whole, and is therefore quite interesting to put into motion. Ideally you will do this procedure on the master node, as it should be 
                                          done, but in theory you could do this from any other node aswell, atleast the first program.

                                          \subsection{MPI Calculation Example - C Version}
                                          This is the aforementioned program that calculates Pi's value utilizing parallel methods to solve it. It is very interesting in the 
                                          sense that it exemplifies exactly what a parallel program should do, and it is also quite simply and therefore allows us to immediately 
                                          gauge and see that our whole tool chain is correctly done. As such this first test is of immense value to us, and will be our true 
                                          starting point for the utilization of our cluster.

                                          \subsubsection{Overview of what will be done}
                                          In order to utilize this code we will need to compile using our new OpenMPI wrapper, so that it has inside of it the proper MPI support
                                          and libraries linked. We will obtain it from its source and then do a quick compilation. Afterwards we will do our first proper test of
                                          the program on both of the machines at once, hopefully with no issues, indicating to us that we've done a good job setting everything 
                                          up, and if for some reason we haven't, we'll probably be able to find out where we screwed up.

                                          \subsubsection{Obtaining the software}
                                          The software in question is part of the examples of an Introduction to Parallel Computing website prepared by 
                                          \textit{computing.llnl.org}, and as such is hosted there. Therefore our first step will be to download the software and place it into 
                                          an appropriate folder. Given the fact that the app is just one source file, we'd do well to simply place the source in our 
                                          \textit{/share/apps} folder, as it really would make no difference. With this in mind, let's do the following:
                                          \begin{lstlisting}
                                          $ cd /share/apps/
                                          $ sudo mkdir -p piexample/noversion/gcc/4.8.5/openmpi/1.10.2
                                          $ cd piexample/noversion/gcc/4.8.5/openmpi/
                                          $ sudo chown pi:pi 1.10.2
                                          $ cd 1.10.2
                                          $ wget computing.llnl.gov/tutorials/mpi/samples/C/mpi_pi_reduce.c
                                          \end{lstlisting}
                                          Now, at this point, we would normally return the folder's permission's to root, however, we will not, as we will go ahead and 
                                          compile it right away right there.

                                          \subsubsection{Preparing to compile}
                                          In order to correctly compile this program you will need to load OpenMPI's module, as it will contain all the necessary include paths 
                                          and library paths to make this work. To do this just run:
                                          \begin{lstlisting}
                                          $ module load openmpi/1.10.2_gcc-4.8.5
                                          \end{lstlisting}
                                          After doing this you should be more than ready to compile the program!

                                          \subsubsection{Compiling the software}
                                          Compiling this program is an extremely simple task, and, additionally, allows you to familiarize yourself with the use of our MPI 
                                          Implementation's GCC wrappers. To do the compilation, all you need to do is run the following command:
                                          \begin{lstlisting}
                                          $ mpicc mpi_pi_reduce.c -o mpi_pi_reduce
                                          \end{lstlisting}
                                          This will give you a file named \textit{mpi\_pi\_reduce} (without the \textit{.c} extension), which is the program we will eventually 
                                          run.

                                          \subsubsection{Testing the software}
                                          Here is where we finally are able to get the gratification of actually running the software on several machines, and this is where 
                                          everything will either go really well or really badly. Thus, with the entireity of our hopes pinned on actually getting this to work 
                                          right, we shall begin by creating a \textit(machinefile), which is a file that OpenMPI will use to know what machines it needs to run 
                                          the program on. As such, your first task is to create a file and open it:
                                          \begin{lstlisting}
                                          $ touch machinefile
                                          $ vi machinefile
                                          \end{lstlisting}
                                          Now, with the file open, you should put the following into it:
                                          \begin{lstlisting}
                                          localhost
                                          compute-0-0
                                          \end{lstlisting}
                                          Essentially, you are telling OpenMPI that you want two processes, one on this machine and one on \textit{compute-0-0}. Now, with that 
                                          done, all that really remains to be is run the program. To do this, go ahead and run the following command\footnote{If you experience
                                              issues with this, substitute \textit{mpirun} with \textit{/share/apps/openmpi/1.10.2/gcc/4.8.5/bin/mpirun} and \textit{machinefile} 
                                              with \textit{/share/apps/piexample/noversion/gcc/4.8.5/openmpi/1.10.2/machinefile}}:
                                              \begin{lstlisting}
                                              $ mpirun -np 2 -machinefile machinefile /share/apps/\
                                                  piexample/noversion/gcc/4.8.5/openmpi/1.10.2/mpi_pi_reduce
                                                  \end{lstlisting}
                                                  After doing this, you should see the program launch and ask you for a password, this, fortunately is something normal, and something 
                                                  we'll eliminate soon, but, for now, just know that you need to input \textit{raspberry} as it is the default password of the 
                                                  \textit{pi} user. Once you do that, the program will continue and you will, hopefully, see the program output it's approximation of Pi
                                                  with no issue. If it does, then everything truly went alright after all.

                                                  \subsection{HPL Performance Benchmark}
                                                  Now comes the time where we get to really see just how much ``oomph'' our system packs ``under the hood''. The HPL Benchmark is the 
                                                  tool that will allow us to do with ease and speed, allowing us to see just how well our system performs under load\footnote{Now, I 
                                                      understand that there are several other benchmarks out there, that, in theory, give a far better and more accurate representation of 
                                                          how well a system will perform, i.e. HPCG, however, as the TOP500 ranking uses HPL, it seems obvious and practical to choose this one 
                                                          over the other.}. This benchmark will also allow us to get a far better feel as to what it will be like to configure, compile and run 
                                                          scientific and distributed software, as they tend to occasionally be quite unwieldy and difficult to put into working order.

                                                          \subsubsection{Overview of what will be done}
                                                          HPL is quite interesting program to set up, as it has been prepared to essentially be as tight and well optimized as possible, with 
                                                          nothing that isn't useful to it within it. Now, in order to get it running, we must realize that the program itself has one truly
                                                          important dependency that has to be built for it, which is some form of a linear algebra library. In this case we have several options
                                                          to choose from, including one the many BLAS implementation or the VSIPL library, however, for our particular case, we will be using 
                                                          OpenBLAS, which, with how we've tested it, has shown the best performance and behavior. The specific version's we'll be using are 2.1 
                                                          for HPL and 0.2.15 for OpenBLAS, as these version have been shown to be quite stable. Now, with this in mind, I'll briefly outline
                                                          the steps we must follow to install HPL and run.
                                                          \begin{itemize}
                                                          \item We must download and obtain the source for both OpenBLAS and HPL.
                                                          \item We must configure, compile and install OpenBLAS 0.2.15.
                                                          \item We must configure, compile and install HPL 2.1.
                                                          \item We must tune and configure HPL's parameters file.
                                                          \item Finally, we must run the application.
                                                          \end{itemize}

                                                          \subsubsection{Obtaining the software}
                                                          This time around, we'll need to download and extract two pieces of software, so go ahead and do the following:
                                                          \begin{lstlisting}
                                                          $ cd
                                                          $ wget http://github.com/xianyi/OpenBLAS/archive/v0.2.15.tar.gz
                                                          $ wget http://www.netlib.org/benchmark/hpl/hpl-2.1.tar.gz
                                                          $ tar xzvf v0.2.15.tar.gz
                                                          $ tar xzvf hpl-2.1.tar.gz
                                                          \end{lstlisting}
                                                          Now that you've finished obtaining the softare, go ahead and head into OpenBLAS's folder:
                                                          \begin{lstlisting}
                                                          $ cd OpenBLAS-0.2.15
                                                          \end{lstlisting}
                                                          With that done, we'll now begin to build it.

                                                          \subsubsection{Compiling and installing OpenBLAS}
                                                          OpenBLAS, unlike most of the other applications we've installed so far, has a rather simple build and
                                                          installation process. This is due to the fact that OpenBLAS has been constructed in a highly portable
                                                          fashion, and, therefore, has great auto-detecting capabilities to make it easily buildable in most 
                                                          environments. Let us then, begin the build process with one very simple command:
                                                          \begin{lstlisting}
                                                          $ make
                                                          \end{lstlisting}
                                                          Once it finishes, you'll need to create the folder structure where you'll install it, and you'll need to have write privilege. So, to 
                                                          do so, do the following:
                                                          \begin{lstlisting}
                                                          $ sudo mkdir -p /share/apps/openblas/0.2.15/gcc/4.8.5/openmpi/1.10.2/
                                                          $ sudo chown pi:pi /share/apps/openblas/0.2.15/gcc/4.8.5/openmpi/1.10.2/
                                                          \end{lstlisting}
                                                          We'll shortly set the permissions back, but for now, do the following:
                                                          \begin{lstlisting}
                                                          $ make install PREFIX=/share/apps/openblas/0.2.15/gcc/4.8.5/openmpi/1.10.2/
                                                          \end{lstlisting}
                                                          After this is done, it is now time to return the folder's permissions to \textit{root}, so that we don't accidentally screw anything 
                                                          up:
                                                          \begin{lstlisting}
                                                          $ sudo chown root:root /share/apps/openblas/0.2.15/gcc/4.8.5/openmpi/1.10.2/
                                                          \end{lstlisting}
                                                          With this now done, you now have the necessary libraries to be able to compile HPL, so that we may run our second and final test.

                                                          \subsubsection{Compiling and Installing HPL}
                                                          HPL, as with most software used on high performance computing clusters, is somewhat tricky to get working, but only somewhat, after a 
                                                          bit, it becomes rather straightforward. In order to correctly configure it, go ahead and move to its folder:
                                                          \begin{lstlisting}
                                                          $ cd ..
                                                          $ cd hpl-2.1
                                                          \end{lstlisting}
                                                          Once you are there, you will need to grab one of the pre-made setup files and modify it so that it is compatible with our system. To do 
                                                          this, copy over one of the files from the \textit{setup} folder, and then rename it to something memorable. After that you must open it 
                                                          and modify it to fit what we will use. Do the following:
                                                          \begin{lstlisting}
                                                          $ cp setup/Make.Linux_PII_CBLAS .
                                                          $ mv Make.Linux_PII_CBLAS Make.PI
                                                          $ vi Make.PI
                                                          \end{lstlisting}
                                                          Now that you've opened it, I will guide you through all the changes you have to make. Line 64 should read:
                                                          \begin{lstlisting}
                                                          ARCH  = PI
                                                          \end{lstlisting}
                                                          Lines 70 through 75 should be:
                                                          \begin{lstlisting}
                                                          TOPdir    = /home/pi/hpl-2.1/
                                                          INCdir    = $(TOPdir)/include
                                                          BINdir    = $(TOPdir)/bin/$(ARCH)
    LIBdir  = $(TOPdir)/lib/$(ARCH)
#
    HPLlib  = $(LIBdir)/hpl.a
    \end{lstlisting}
    Lines 84 tom 86 should be:
    \begin{lstlisting}
    MPdir   = /share/apps/openmpi/1.10.2/gcc/4.8.5/
    MPinc   = -I$(MPdir)/include
    MPlib   = $(MPdir)/lib/libmpi.so
    \end{lstlisting}
    Lines 95 to 97 should be:
    \begin{lstlisting}
    LAdir   = /share/apps/openblas/0.2.15/gcc/4.8.5/openmpi/1.10.2/lib/
    LAinc   =
    LAlib   = $(LAdir)/libopenblas.a
    \end{lstlisting}
    Lines 169 to 171 should be:
    \begin{lstlisting}
    CC  = /share/apps/openblas/0.2.15/gcc/4.8.5/openmpi/1.10.2/bin/mpicc
    CCNOOPT = $(HPL_DEFS)
    CCFLAGS = $(HPL_DEFS) -fomit-frame-pointer -O3 -funroll-loops
    \end{lstlisting}
    Line 176 should be:
    \begin{lstlisting}
    LINKER  = /share/apps/openblas/0.2.15/gcc/4.8.5/openmpi/1.10.2/bin/mpicc
    \end{lstlisting}
    Go ahead and save the file. Once you've done that, then you'll be ready to build the program itself. This 
    can be done with the following commands:
    \begin{lstlisting}
    $ make arch=PI
    \end{lstlisting}
    Once you are done with this, you will be able to find and executable within a newly created folder. It is
    this very file that'll be the one that will allow us to run all of our desired tests. Please verify that an 
    executable named \textit{xhpl} is present under the folder \textit{bin/PI/}. To do this, do the following:
    \begin{lstlisting}
    $ ls bin/PI/
    \end{lstlisting}
    You should see, within, the file previously mentioned. Additionally you should find a file named 
    \textit{HPL.dat}, which will be the file we'll be using to do our runtime configuration for \textit{HPL}. 
    If you successfully, found the files, it is now time to perform the install. Do the following commands to
    do so:
    \begin{lstlisting}
    $ sudo mkdir -p /share/apps/hpl/2.1/gcc/4.8.5/openmpi/1.10.2/
    $ sudo chown pi:pi /share/apps/hpl/2.1/gcc/4.8.5/openmpi/1.10.2/
    $ mv bin/PI/{xhpl,HPL.dat} /share/apps/hpl/2.1/gcc/4.8.5/openmpi/1.10.2/
    $ sudo chown -R root:root /share/apps/hpl/2.1/gcc/4.8.5/openmpi/1.10.2/
    \end{lstlisting}
    Once you've done this, you'll now have a working implementation of \textit{HPL} that you can run.

    \subsubsection{Testing the software}
    Now that all, in terms of building and installing, is done, it is time to finally test the software, to 
    guarantee that our system is now stable and in working order. To achieve this goal, we will follow a path 
    that is quite similar to what we did, not too long ago, with our Pi calculator. The steps we will follow 
    are quite simple, and not difficult in the slightest, as it all pretty much is just a rehash of what we've 
    done before, with some slight changes. To begin, we must edit the HPL.dat file, so, go ahead and open it:
    \begin{lstlisting}
    $ cd /share/apps/hpl/2.1/gcc/4.8.5/openmpi/1.10.2/
    $ sudo vi HPL.dat
    \end{lstlisting}
    Now that you've opened it, you'll notice that this file is filled with configurations. Most of the 
    parameters, in theory can be left as is, though with, some, notably the N size, which determines how much 
    ram you can utilize at once, and the NB, which determines how the matrix will be split, are perfect 
    examples of parameters that need to be changed. In addition to those, are the P and Q lines, as these 
    indicate how your cluster is laid out. And, since ours is only 2 machines with one core each, we can only 
    really run the setup as it is shown in the file. With all of this in mind, I will now provide you with the configuration I found optimal in this setup for both of our machines:
    \begin{lstlisting}
    HPLinpack benchmark input file
    Innovative Computing Laboratory, University of Tennessee
    HPL.out      output file name (if any) 
    6            device out (6=stdout,7=stderr,file)
    1            # of problems sizes (N)
    9984         Ns
    1            # of NBs
    192           NBs
    0            PMAP process mapping (0=Row-,1=Column-major)
    1            # of process grids (P x Q)
    1            Ps
    2            Qs
    16.0         threshold
    1            # of panel fact
    2            PFACTs (0=left, 1=Crout, 2=Right)
    1            # of recursive stopping criterium
    4            NBMINs (>= 1)
    1            # of panels in recursion
    2            NDIVs
    1            # of recursive panel fact.
    1            RFACTs (0=left, 1=Crout, 2=Right)
    1            # of broadcast
    1            BCASTs (0=1rg,1=1rM,2=2rg,3=2rM,4=Lng,5=LnM)
    1            # of lookahead depth
    1            DEPTHs (>=0)
    2            SWAP (0=bin-exch,1=long,2=mix)
    64           swapping threshold
    0            L1 in (0=transposed,1=no-transposed) form
    0            U  in (0=transposed,1=no-transposed) form
    1            Equilibration (0=no,1=yes)
    8            memory alignment in double (> 0)
    \end{lstlisting}
    Now, armed with this, we are only one step away from actually being able to run our test, and, without 
    further ado, here's what you need to do to run it:
    \begin{lstlisting}
    $ sudo cp /share/apps/piexample/noversion/gcc/4.8.5\
        /openmpi/1.10.2/machinefile .
        $ mpirun -np 2 -machinefile machinefile xhpl
        \end{lstlisting}
        And with that, the test should be underway and you'll get a real world performance report when it finishes.
        Hopefully you'll be delighted to see that the cluster works quite well, and delivers quite good performance. With this in mind, don't 
        forget to save that HPL.dat configuration, as you may come back to tweak it to learn more about optimizing parallel apps.

        \section{Distributed John the Ripper}
        Openwall's John the ripper has, for quite some time been a wonderful application for anybody interested in the world of cyber security, 
        as it is a powerful and flexible program for cracking hashes and retrieving the passwords and access codes that lay within. However, a 
        machine's performance has always been a deciding factor as to how effective John the ripper is at cracking the hashes it's fed. With 
        this in mind, it should be no wonder that perhaps if two machines were to work in parallel, they could surely crack the passwords far 
        easier than if they weren't. Now, john the ripper is released in three versions, a very stable and supported ``Pro'' version, a 
        standard ``free'' version, and the most interesting of them all, the community enhanced ``jumbo'' version, where members of John the 
        Ripper's community have taken the time to add many of the features they feel are lacking with the official john the ripper version. 
        That version, then, will be the one we will be using, as it'll afford us the greatest flexibility and greatest amount of features. 
        Additionally, it is easier to install and deploy on a system than its free (and official\footnote{Nonetheless, they officially endorse 
                and provide it on John the Ripper's website, and as such, really seems like the best alternative to use, shy of actually buying the 
                official version.}) counterpart. With all of this now absorbed into our minds, let us now begin to set it up.

        \subsection{Overview of what's been done and what will be done}
        As you might probably be imagining, we've had to wade through an enormous amount of work to get here, and, well, it must be explained. 
        First off, the latest John the Ripper ``jumbo'' contains quite a large body of work done by community programmers that decided to use 
        some of the new features that were introduced for C++14, and unfortunately, compilers that didn't support this standard were simply 
        unable to produce a working John the Ripper executable. Given the fact that the Raspberry Pi, by default, provides a very old version
        of GCC, it is no wonder then that it was necessary to move onward to a newer version of GCC. This had the fantastic side-effect of 
        allowing me to introduce a large part of what needs to be done to configure and get a cluster in working order, at least at the very 
        beginning, and it allowed me to present how and why things need to be done in order to use a real world application on a distributed 
        system. Many of the things we've done, simply are the way they are so we can actually use John the Ripper, for example, in order to 
        have a distributed John the Ripper work, you need to have the hash files somewhere where they can be easily found and read by all of 
        the systems, inherently necessitating an NFS. This, basically, is what has allowed us to reach this point, where we are nearly ready to 
        run the application that is the true focus of this work, a distributed version of John the Ripper. I will now present and outline, of 
        the work that lies ahead, and afterwards illustrate what the final product of all of this work will be. To build and prepare the 
        application for use, we will need to follow these steps:
        \begin{itemize}
        \item We will need to download the latest stable version of John the Ripper's ``jumbo'' community maintained version.
        \item We will then need to extract the software to a suitable location.
        \item Once that is done, we'll proceed to configure the software, so that it may be built according to our specific needs.
        \item We'll compile the software package and subsequently install it.
        \item We'll prepare a suitable set of hashes to be cracked.
        \end{itemize}
        After all of this is said and done, we'll finally have a tried and true cluster of our own, with a near infinite amount of 
        possibilities lying ahead, just waiting to be undertaken.

        \subsection{Obtaining the software}
        Obtaining the software is rather simple, and really, given the fact that there are several options, any of them are perfectly fine. We
        will be using Openwall's official download link. Once we have it, we'll go ahead and extract it too. To obtain the source, run the 
        following commmands:
        \begin{lstlisting}
        $ cd
        $ wget http://www.openwall.com/john/j/john-1.8.0-jumbo-1.tar.gz
        $ tar xzvf john-1.8.0-jumbo-1.tar.gz
        $ cd john-1.8.0-jumbo-1/src
        \end{lstlisting}
        Once you're there you should see a myriad of files, of all types, but, more important than anything else, you must be sure that there 
        is a working configure file.

        \subsection{Configuring the software}
        John the Ripper, by default does not include a \textit{configure} built in, making its build process more difficult and involved than 
        it really should be (although it is simple). Fortunately for us, the ``jumbo'' version has a quite robust and well designed 
        \textit{configure script} which makes the build process rather straightforward. Just make sure that you have your OpenMPI module loaded 
        once you begin configuring. To configure this program, run the following command:
        \begin{lstlisting}
        $ ./configure --prefix=/share/apps/johntheripper/\
                               1.8.0-jumbo-1/gcc/4.8.5/openmpi/1.10.2/ --enable-mpi\
                               --disable-cuda --disable-openmp --build=arm-linux-gnueabihf
                               \end{lstlisting}
                               Once it finishes, we should create the folder structure for it and get it ready for the installation with the following commands:
                               \begin{lstlisting}
                               $ sudo mkdir -p /share/apps/johntheripper/1.8.0-jumbo-1\
                                   /gcc/4.8.5/openmpi/1.10.2
                                   $ sudo chown pi:pi /share/apps/johntheripper/1.8.0-jumbo-1\
                                       /gcc/4.8.5/openmpi/1.10.2
                                       \end{lstlisting}
                                       Now, with that done, it is time to begin compiling the software.

                                       \subsection{Compiling and installing the software}
                                       Compiling john the ripper, thankfully, is quite straightforward, and therefore really doesn't require much introduction. To begin 
                                       compiling, run these commands:
                                       \begin{lstlisting}
                                       $ make
                                       $ make install
                                       \end{lstlisting}
                                       After this, in theory, john the ripper is done and installed and ready to use. So, with that in mind, I think it would make our lives 
                                       easier to be able to call up John the Ripper and any of his abilities on demand, and, therfore, it makes sense for us to now go ahead 
                                       and create a simple module for it.

                                       \subsection{Preparing a modulefile for John the Ripper}
                                       Without overextending myself, let me just say that, as before, having a module for john the ripper makes life a bit easier when working 
                                       on the fly. Here is one module that would probably allow you to work with john quite well. First, let's create and edit our modulefile:
                                       \begin{lstlisting}
                                       $ cd /share/apps/modules/3.2.10/modulefiles/
                                       $ sudo mkdir johntheripper
                                       $ sudo vi johntheripper/1.8.0-jumbo-1_gcc-4.8.5_openmpi-1.10.2
                                       \end{lstlisting}
                                       Now that it's open, go ahead and edit it with the following info:
                                       \begin{lstlisting}
#%Module1.0#####################################################
##
## Modulefile for John the Ripper 1.8.0-jumbo-1
##
##
##
                                       proc ModulesHelp { } {
                                                   global version modroot

                                                               puts stderr "Prepares the environment for John the Ripper"
                                       }

module-whatis "Modulefile for John the Ripper 1.8.0-jumbo-1"

module load openmpi/1.10.2_gcc-4.8.5

set     topdir  /share/apps/johntheripper/1.8.0-jumbo-1\
            /gcc/4.8.5/openmpi/1.10.2
            set     version 1.8.0-jumbo-1
            set     app John the Ripper

            prepend-path    PATH      $topdir/run
            \end{lstlisting}
            This should be all you need to get it up and running. At this point you really are ready to begin testing John the Ripper.

            \subsection{Preparing a set of hashes for testing}
            Now that you've produced a working executable, it is time to begin assembling a suitable set of hashes to 
            crack. Personally, as this is more of test case than anything else, I believe it would be worth it to make 
            it as simple as can be, as such, my suggestion is to create just 3 users and create low character count 
            passwords, so you can tell that the software works, without committing to several hours of work trying to 
            crack the passwords you've created. To do this, run these commands to generate the users:
            \begin{lstlisting}
            $ sudo useradd user1 -p pw1
            $ sudo useradd user2 -p pw2
            $ sudo useradd user3 -p pw3
            \end{lstlisting}
            Now that you've done this, the system has 3 new users that are ready to crack. Unfortunately, we can't
            crack them quite yet. Modern linux systems have the user info carefully split among two files, 
            \textit{/etc/passwd} and \textit{/etc/shadow}, and, in order to be able to crack it, you must use a utility 
            provided by John the Ripper named \textit{unshadow} to combine them into a single file. In order to do 
            this, go ahead and run these commands: 
            \begin{lstlisting}
            $ sudo su
# module load johntheripper/1.8.0-jumbo-1_gcc-4.8.5_openmpi-1.10.2
# unshadow /etc/passwd /etc/shadow > passdb.txt
# chown pi:pi passdb.txt
# mkdir /share/files
# mv passdb.txt /share/files/.
# cp /share/apps/piexample/noversion/gcc/4.8.5\
                /openmpi/1.10.2/machinefile /share/files/.
# logout
            $ cd /share/files
            $ module johntheripper/1.8.0-jumbo-1_gcc-4.8.5_openmpi-1.10.2
            \end{lstlisting}
            Now, finally, we have a file we can actually use, and now, finally, it is time to put our cluster to use.

            \subsection{Cracking the passwords}
            After all, we are now, thankfully, one command from completing the goals we've set out for ourselves, and 
            that, in it of itself is more than enough reason to celebrate. Now, basking in the greatness of our 
            accomplishments, let us now run this command:
            \begin{lstlisting}
            $ mpirun -np 2 -machinefile machinefile john passdb.txt
            \end{lstlisting}
            Once it begins to churn through the passwords, it'll in show you that it has successfully cracked three 
            passwords, and as such, it is, coupled to our earlier tests, absolute proof that you've constructed and put 
            into working order a fantastic little cluster.

            \subsection{Performance}
            Now, as a small though on the matter, during my testing I found that the real world performance of John the Ripper was highly dependent 
            on how well the network itself performed. However, if used on a network switch with reasonable speeds, there was a slight improvement 
            in runtime cracking up to three passwords at a time. This was on the order of around a couple of minutes. However, once I did my 
            testing with a switch from 1990, I found myself distraught by waiting over an hour for it to finish, instead of mere minutes on a 
            single machine. There certainly is a great deal of work to be done yet experimenting and investigating, so this is still an interesting
            open question.

            \section{Thoughts for future expansion}
            Here at the end of this journey comes a time where now you must decide what you want to use this cluster 
            for, in this case, one awesome usage case would be to have it be a small and almost non power consuming 
            utility to check whether all the users on a given cluster are using safe passwords. But, to be completely 
            frank and sincere, I believe that the possibilities are truly endless, and therefore the onus is now on you 
            to demonstrate your abilities by proving that you are more than up to the challenge of continuing this 
            cluster. If nothing else, this is a fantastic educational tool that can be used to teach a new generation 
            of highly involved and competent system administrators.

            \section{Conclusion}
            Having said just about everything, now comes the time to truly reflect on what has been done. Though at 
            times it may feel that this project leaves you wandering in the dark, an overwhelming sense of purpose 
            overcomes you once you hit the first successes, and that is the true fuel and power that this tool has. The 
            motivation of slowly creating something new out of what seems something completely detached to the world of 
            high performance computing is perhaps the most powerful that can be given to a student.

            This document's purpose is no other than to serve a small guiding light into the world of high performance 
            computing, and, hopefully, it has lived up to be what it set out to be.

            \newpage
            \section{References}
            \begin{itemize}
            \item D'Amore, M., Baggio, R., \& Valdani, E. (2014). A Practical Approach to Big Data in Tourism: A Low Cost Raspberry Pi Cluster. Information and Communication Technologies in Tourism 2015, 169-181.
            \item Furlani, J. L. (1991). Modules: Providing a Flexible User Environment. Proceedings of the Fifth Large Installation Systems Administration Conference (LISA V), 141-152.
            \item GNU MP 6.1.0. (n.d.). Retrieved February 04, 2016, from https://gmplib.org/manual/index.html
            \item GNU MPFR 3.1.3. (n.d.). Retrieved February 04, 2016, from http://www.mpfr.org/mpfr-current/mpfr.html
            \item Gite, V. (2007, December 18). Ubuntu Linux NFS Server installation and Configuration. Retrieved February 04, 2016, from http://www.cyberciti.biz/faq/how-to-ubuntu-nfs-server-configuration-howto/
            \item Gowtham S. (2007, July 2). HPL Benchmark For Single Processor Machines. Retrieved February 04, 2016, from http://sgowtham.com/journal/hpl-benchmark-for-single-processor-machines/
            \item Gowtham S. (2012, February 27). Rocks 5.4.2  HPL 2.0 benchmark with GCC 4.1.2. Retrieved February 04, 2016, from http://sgowtham.com/journal/hpl-2-0-benchmark-with-gcc-4-1-2-on-rocks-5-4-2/
            \item Html Documentation GNU MPC 1.0.3. (n.d.). Retrieved February 04, 2016, from http://www.multiprecision.org/index.php?prog=mpc
            \item Installing GCC: Configuration. (n.d.). Retrieved February 04, 2016, from https://gcc.gnu.org/install/configure.html
            \item Kiepert, J. (2013, May 22). Creating a Raspberry Pi-Based Beowulf Cluster. Retrieved February 4, 2016, from http://coen.boisestate.edu/ece/files/2013/05-
            /Creating.a.Raspberry.Pi-Based.Beowulf.Cluster\_v2.pdf
            \item Leonard, P. (2012, June 18). Parallel Processing on the Pi (Bramble). Retrieved February 04, 2016, from http://westcoastlabs.blogspot.co.uk/2012/06/parallel-processing-on-pi-bramble.html
            \item Modules Software Environment. (n.d.). Retrieved February 04, 2016, from https://www.nersc.gov/users/software/nersc-user-environment/modules/
            \item Network configuration. (n.d.). Retrieved February 04, 2016, from https://wiki.archlinux.org
            /index.php/Network\_configuration
            \item OpenBLAS User Manual. (n.d.). Retrieved February 04, 2016, from https://github.com/xianyi/OpenBLAS/wiki/User-Manual
            \item OpenMPI Project. (n.d.). FAQ: Building Open MPI. Retrieved February 04, 2016, from https://www.open-mpi.org/faq/?category=building
            \item Parallel and distributed processing with John the Ripper. (n.d.). Retrieved February 04, 2016, from http://openwall.info/wiki/john/parallelization
            \item RPi Distributions. (n.d.). Retrieved February 04, 2016, from 
            http://elinux.org/RPi\_Distributions
            \item Raspberry Pi Foundation (n.d.). What is a Raspberry Pi? Retrieved February 4, 2016, from https://www.raspberrypi.org/help/what-is-a-raspberry-pi/
            \item Raspbian Team. (n.d.). Raspbian FAQ. Retrieved February 04, 2016, from https://www.raspbian.org/RaspbianFAQ
            \item Redmond, E. (2012). Building a Riak Cluster on Raspberry Pi. Retrieved February 04, 2016, from http://basho.com/posts/technical/building-a-riak-cluster-on-raspberry-pi/
            \item Whitney, E., \& Sprague, M. (2001). Drag your design environment kicking and screaming into the '90s with Modules! Synopsys Users' Group. Retrieved February 4, 2016, from http://modules.sourceforge.net/docs/MC2\_whitney\_paper.pdf

            \end{itemize}

            \end{document}

